---
title: "CLI"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F)
library(tidyverse)
library(readxl)
library(DT)
#install.packages("remotes")
#remotes::install_github("rstudio/fontawesome")
library(fontawesome)

f <- "./static/commands.xlsx"
```

---

# Operating Systems

## General

[SS64](https://ss64.com/)<br>
Reference guide containing syntax and examples for the most prevalent computing commands (Database and Operating System).<br>

### SSH

`ssh <username>@<host ip>`

## Mac

### Movie > Gif

`ffmpeg -i file.mov -s 600x400 -pix_fmt rgb24 -r 20 -f gif - | gifsicle --optimize=3 --delay=3 > file.gif`

### .zshrc
```{}
export GPG_TTY=$TTY

# Enable Powerlevel10k instant prompt. Should stay close to the top of ~/.zshrc.
# Initialization code that may require console input (password prompts, [y/n]
# confirmations, etc.) must go above this block; everything else may go below.
if [[ -r "${XDG_CACHE_HOME:-$HOME/.cache}/p10k-instant-prompt-${(%):-%n}.zsh" ]]; then
  source "${XDG_CACHE_HOME:-$HOME/.cache}/p10k-instant-prompt-${(%):-%n}.zsh"
fi

# If you come from bash you might have to change your $PATH.
# export PATH=$HOME/bin:/usr/local/bin:$PATH

# Path to your oh-my-zsh installation.
export ZSH="/Users/cgrant/.oh-my-zsh"

# Set name of the theme to load --- if set to "random", it will
# load a random theme each time oh-my-zsh is loaded, in which case,
# to know which specific one was loaded, run: echo $RANDOM_THEME
# See https://github.com/ohmyzsh/ohmyzsh/wiki/Themes
ZSH_THEME="powerlevel10k/powerlevel10k"

# Set list of themes to pick from when loading at random
# Setting this variable when ZSH_THEME=random will cause zsh to load
# a theme from this variable instead of looking in $ZSH/themes/
# If set to an empty array, this variable will have no effect.
# ZSH_THEME_RANDOM_CANDIDATES=( "robbyrussell" "agnoster" )

# Uncomment the following line to use case-sensitive completion.
CASE_SENSITIVE="false"

# Uncomment the following line to use hyphen-insensitive completion.
# Case-sensitive completion must be off. _ and - will be interchangeable.
# HYPHEN_INSENSITIVE="true"

# Uncomment one of the following lines to change the auto-update behavior
# zstyle ':omz:update' mode disabled  # disable automatic updates
# zstyle ':omz:update' mode auto      # update automatically without asking
zstyle ':omz:update' mode reminder  # just remind me to update when it's time

# Uncomment the following line to change how often to auto-update (in days).
zstyle ':omz:update' frequency 13

# Uncomment the following line if pasting URLs and other text is messed up.
# DISABLE_MAGIC_FUNCTIONS="true"

# Uncomment the following line to disable colors in ls.
# DISABLE_LS_COLORS="true"

# Uncomment the following line to disable auto-setting terminal title.
# DISABLE_AUTO_TITLE="true"

# Uncomment the following line to enable command auto-correction.
# ENABLE_CORRECTION="true"

# Uncomment the following line to display red dots whilst waiting for completion.
# You can also set it to another string to have that shown instead of the default red dots.
# e.g. COMPLETION_WAITING_DOTS="%F{yellow}waiting...%f"
# Caution: this setting can cause issues with multiline prompts in zsh < 5.7.1 (see #5765)
# COMPLETION_WAITING_DOTS="true"

# Uncomment the following line if you want to disable marking untracked files
# under VCS as dirty. This makes repository status check for large repositories
# much, much faster.
# DISABLE_UNTRACKED_FILES_DIRTY="true"

# Uncomment the following line if you want to change the command execution time
# stamp shown in the history command output.
# You can set one of the optional three formats:
# "mm/dd/yyyy"|"dd.mm.yyyy"|"yyyy-mm-dd"
# or set a custom format using the strftime function format specifications,
# see 'man strftime' for details.
# HIST_STAMPS="mm/dd/yyyy"

# Would you like to use another custom folder than $ZSH/custom?
# ZSH_CUSTOM=/path/to/new-custom-folder

# Which plugins would you like to load?
# Standard plugins can be found in $ZSH/plugins/
# Custom plugins may be added to $ZSH_CUSTOM/plugins/
# Example format: plugins=(rails git textmate ruby lighthouse)
# Add wisely, as too many plugins slow down shell startup.
plugins=(
  git
  docker
  zsh-syntax-highlighting
  zsh-autosuggestions
  web-search
  copypath
  copyfile
  copybuffer
  macos
  1password
)

source $ZSH/oh-my-zsh.sh

# User configuration

# export MANPATH="/usr/local/man:$MANPATH"

# You may need to manually set your language environment
# export LANG=en_US.UTF-8

# Preferred editor for local and remote sessions
# if [[ -n $SSH_CONNECTION ]]; then
#   export EDITOR='vim'
# else
#   export EDITOR='mvim'
# fi

# Compilation flags
# export ARCHFLAGS="-arch x86_64"

# Set personal aliases, overriding those provided by oh-my-zsh libs,
# plugins, and themes. Aliases can be placed here, though oh-my-zsh
# users are encouraged to define aliases within the ZSH_CUSTOM folder.
# For a full list of active aliases, run `alias`.
#
# Example aliases
# alias zshconfig="mate ~/.zshrc"
# alias ohmyzsh="mate ~/.oh-my-zsh"

# STANDARD FUNCTION MODIFICATIONS
alias cat='ccat'
alias l='exa -1 -g -h -l --git --group-directories-first --icons --sort=Name --time=modified'
alias la='exa -1 -g -h -l --git --group-directories-first --icons --sort=Name --time=modified -a'
# SHORTCUTS
alias c='clear'
alias h='history'
alias r='source ~/.zshrc'
alias speed='speedtest-cli --simple'
# HOMEBREW
alias casks='brew cu -a --no-brew-update -v'
alias brewup='brew update; brew upgrade; brew cu -a --no-brew-update -v; brew cleanup; brew doctor'
# PATHS
alias geo='cd /Users/cgrant/OneDrive/GitHub/geocoug'
alias cloud='cd "/Users/cgrant/Library/Mobile Documents/com~apple~CloudDocs"'
alias one='cd /Users/cgrant/OneDrive'
# OWNERSHIP
alias my='sudo chown -R `id -u`'
# GIT
alias gaa='git add --all'
# DOCKER / ENVIRONMENTS
alias repo='docker run --rm -it -v /Users/cgrant/Library/CloudStorage/OneDrive-Personal/GitHub/github.txt:/app/github.txt create-repo python create_repo.py -f /app/github.txt -r'
alias lab='docker run -it --rm -p 8888:8888 -v $(PWD):/home/jovyan jupyter-lab'
alias dev='source ~/venvs/dev/bin/activate'
alias pyenv='/opt/homebrew/opt/python@3.10/Frameworks/Python.framework/Versions/3.10/bin/python3.10 -m venv .venv && source ./.venv/bin/activate && python -m pip install --no-cache-dir pre_commit flake8 black isort autoflake bandit mypy > /dev/null 2>&1 && python -m pip freeze > requirements.txt > /dev/null 2>&1 && code .'


# To customize prompt, run `p10k configure` or edit ~/.p10k.zsh.
[[ ! -f ~/.p10k.zsh ]] || source ~/.p10k.zsh


export PATH="/Users/cgrant/bin:$PATH"

export PATH="/opt/homebrew/opt/curl/bin:$PATH"
```


## Windows

### CMD
[Microsoft Docs](https://docs.microsoft.com/en-us/windows-server/administration/windows-commands/cmd)<br>
Command line syntax

**Syntax**: `cmd [/c|/k] [/s] [/q] [/d] [/a|/u] [/t:{<b><f> | <f>}] [/e:{on | off}] [/f:{on | off}] [/v:{on | off}] [<string>]`

**Multiple commands**: `"<command1>&&<command2>&&<command3>"`

#### Parameters
```{r cmd, echo=F}
t1 <- read_excel(f, sheet="windows-commands")
# Show all rows in table (no pagination)
datatable(t1, options=list(paging=FALSE), rownames=F)
```

### Powershell 

<div style="display:inline;">`r fa("hard-hat")` `r fa("hammer")` Under construction...</div>

---

## Linux

### Commands
```{r}
t2 <- read_excel(f, sheet="linux-commands")
# Show all rows in table (no pagination)
datatable(t2, options=list(paging=FALSE), rownames=F)
```

### File System
```{r, echo=F}
t3 <- read_excel(f, sheet="linux-filesystem")
# Show all rows in table (no pagination)
datatable(t3, options=list(paging=FALSE), rownames=F)
```

### Permissions

General syntax: `_rwxrwxrwx 1 owner group`<br>

`_ | rwx | rwx | rwx` <Special><Owner><Group><All Users>

`r` = Read<br>
`w` = Write<br>
`x` = Execute

`chown` = Change ownership<br>
`chmod` = Change permissions<br>

`sudo chown <user>:<group> <file>`

`sudo chown 664 <file>`

### Package Manager
```{r, echo=F}
t4 <- read_excel(f, sheet="linux-packages")
# Show all rows in table (no pagination)
datatable(t4, options=list(paging=FALSE), rownames=F)
```


### Server Setup
Initial Linux server setup on Digital Ocean

1) Login to server as root
2) `$ adduser <username>`
  a. Creates new admin user
  Add user to sudo group: `usermod -aG sudo <username>`
3) `$ exit`
4) Reload SSH connection
5) Login as new user
6) `$ sudo vi /etc/ssh/sshd_config`
  a. Configuration fo rputty/incoming SSH connections
  b. Change `PermitRootLogin` to `no`
7) `$ sudo reboot`
8) On boot, can see linux kernel used for current build/image
  a. `GNU/Linux 5.4.0-51-generic`
9) `$ apt list --upgradable`
10) `$ sudo apt upgrade`
11) `$ uname -a`
  a. Shows system info
  b. Even though you updated the kernel, need to reboot to take effect
12) `$ sudo reboot`
13) `$ sudo apt update`
	a. Just to be sure 
14) `$ sudo apt upgrade`
15) `$ ufw`
	a. *Uncomplicated firewall
16) `$ sudo ufw status verbose`
	a. Shows firewall status
17) `$ sudo ufw help`
18) `$ sudo ufw allow OpenSSH`
	a. Will allow opening of ports
19) `$ sudo ufw enable`
	a. Turns on firewall
20) `$ sudo ufw status verbose`
  a. Shows status of firewall/open ports
21) Good time for snapshot backup
	a. Has to be shutdown for backup/snapshot
22) `$ sudo shutdown -h now`
You have to go to DigitalOcean and reboot

---

# Docker

## Resources

- [CheatSheet](./static/Docker-CheatSheet.pdf)
- [Convert Docker command to docker-compose.yml](https://www.composerize.com/)

## Image vs. Container

**Image** - Application we want to run

**Container** - Instance of that image running as a process

---

## Docker Basics

### Create an Nginx container

`docker run -p 80:80 -d --name webhost nginx`

1. Downloads Nginx from Docker Hub
1. Starts new container from that image
1. Opened port 80 on host IP
1. Routes port 80 traffic to the container IP, port 80
1. View container at [http://localhost:80](http://localhost:80)

### Other examples

`docker run -p 80:80 -d --name nginx nginx`

`docker run -p 8080:80 -d --name httpd httpd`

`docker run -p 3306:3306 --platform linux/amd64 -d --name mysql -e MYSQL_RANDOM_ROOT_PASSWORD=true mysql`

Create a JupyterLab instance and attach your current directory as a volume: `docker run -it --rm -p 8888:8888 -v $(PWD):/home/jovyan jupyter/pyspark-notebook`

### Processes and configurations

Check processes running inside a container: `docker top <container>`

Container configuration: `docker <container> inspect`

Check container stats (memory, cpu, network): `docker stats <container>`

### Getting a shell inside containers

Start a new container interactively: `docker run -it <container>`

Run commands in existing container: `docker exec -it <container>`

#### Example: Start a container interactively and launch bash within it

1. Start container and launch bash: `docker run -it --name ubuntu ubuntu bash`
1. Run some bash command: `apt-get install -y curl`
1. Exit the container: `exit`
1. Start and re-enter the container: `docker start -ai ubuntu`

#### Example: Launch shell in running container

`docker exec -it <container> bash`

### Pull an image from docker hub

`docker pull <imagename>`

---

## Docker Networks

-   Each container is connected to a private virtual network (called "bridge").
-   Each virtual network routes through NAT firewall on host IP.
-   All containers on a virtual network can talk to each other without `-p`
-   **_Best practice_**: Create a new virtual network for each app.
-   You can skip virtual networks and use the host IP (`--net=host`).

Get container IP: `docker inspect --format '{{ .NetworkSettings.IPAddress }}' <container>`

### Publishing (#:#)

example: 8080:80

**left number**: published/host port

**right number**: listening/container port

_Traffic passing through port 8080 on the HOST will be directed to port 80 on the container._

### DNS

Docker uses container names as host names.

Dont rely on IPs for inter-communication.

**_Best Practice_** Always use custom networks.

#### Assignment

Check different curl versions within current versions of Ubuntu and CentOS.

Run "curl --version" on both operating systems.

##### Steps

**ubuntu**: `apt-get update && apt-get install curl`

**centos**: `yum update curl`

Then...

`curl --version`

Also:

Check out command `docker --rm`

## Dockerfiles

Recipe for creating images 

Each Dockerfile stanza such as "RUN", "CMD", etc. are stored as a single image layer. Docker caches each layer by giving it a unique SHA (hash), so whenever the image is (re)built, it can check to see if a layer has changed, and if not, it will use the cached layer.

Docker builds images top down, so it is best practice to structure the Dockerfile in such a way that lines which will change the most are at the bottom, and lines that will change the least are at the top. If a line is changed (ie. source code changes) Docker will rebuild that line, and thus each line after that will also need to be rebuilt.

## Keeping the Docker system clean

`docker system prune`
- all stopped containers
- all networks not used by at least one container
- all dangling images
- all dangling build cache

## Volumes an Bind Mounts

**Volumes** - Special location outside of container UFS

**Bind Mounts** - Link container path to host path

Build an image and ***named*** volume (persistent): `docker run -d --name mysql -e MYSQL_ALLOW_EMPTY_PASSWORD=True -v mysql:/var/lib/mysql --platform linux/amd64 mysql`

---

# Git

- [https://product.hubspot.com/blog/git-and-github-tutorial-for-beginners](https://product.hubspot.com/blog/git-and-github-tutorial-for-beginners)
- [https://docs.gitlab.com/ee/gitlab-basics/start-using-git.html](https://docs.gitlab.com/ee/gitlab-basics/start-using-git.html)

## .gitconfig
```{}
# This is Git's per-user configuration file.
[init]
	defaultBranch = main
[user]
  useConfigOnly = true
	name = Caleb Grant
  email = 
  signingKey = 
[core]
	excludesfile = /Users/cgrant/.gitignore_global
	editor = vim
[difftool "sourcetree"]
	cmd = opendiff \"$LOCAL\" \"$REMOTE\"
	path =
[mergetool "sourcetree"]
	cmd = /Applications/Sourcetree.app/Contents/Resources/opendiff-w.sh \"$LOCAL\" \"$REMOTE\" -ancestor \"$BASE\" -merge \"$MERGED\"
	trustExitCode = true
[commit]
	template = /Users/cgrant/.stCommitMsg
	gpgsign = true
[gpg]
	program = gpg
[filter "lfs"]
	process = git-lfs filter-process
	required = true
	clean = git-lfs clean -- %f
	smudge = git-lfs smudge -- %f
[credential]
	helper = osxkeychain
```




## Initialize
1) Launch Git Bash
2) Navigate to project directory
3) initialize git repository in the folder root: `git init`
4) create new file in directory: `touch filename.extension`
5) list files in root: `ls`
6) check which files git recognizes: `git status`


## Staging
A commit is a record of what files you have changed since the last time you made a commit. 
Essentially, you make changes to your repo (for example, adding a file or modifying one) 
and then tell git to put those files into a commit.
Commits make up the essence of your project and allow you to go back to the state of a project at any point.

So, how do you tell git which files to put into a commit? 
This is where the staging environment or index come in. 
When you make changes to your repo, git notices that a file has changed but 
won't do anything with it (like adding it in a commit).

To add a file to a commit, you first need to add it to the staging environment. 
To do this, you can use the `git add <filename>` command.

Once you've used the git add command to add all the files you want to the staging environment, 
you can then tell git to package them into a commit using the git commit command. 
Note: The staging environment, also called 'staging', is the new preferred term for this, 
but you can also see it referred to as the 'index'.

1) Add files to the staging environment: `git add filename.extension`
2) Check staging environment for new files: `git status`


## Commit Locally
`git commit -m "Your message about the commit"`


## Branches
Say you want to make a new feature but are worried about making changes to the main project 
while developing the feature. This is where git branches come in. 

Branches allow you to move back and forth between 'states' of a project. 
For instance, if you want to add a new page to your website you can create a new branch just 
for that page without affecting the main part of the project. Once you're done with the page, 
you can merge your changes from your branch into the master branch. 
When you create a new branch, Git keeps track of which commit your branch 'branched' off of, 
so it knows the history behind all the files. 

1) `git checkout -b <my branch name>`
2) Show list of branches: `git branch`


## Commit to Github
1) Create new repo on GitHub
2) `git remote add origin <url produced on github for new repo>`
3) `git push -u origin [master/main]`


## Push a Branch to Github
`git push origin <my-new-branch>`

You might be wondering what that "origin" word means in the command above. 
What happens is that when you clone a remote repository to your local machine, git creates an alias for you. 
In nearly all cases this alias is called "origin." It's essentially shorthand for the remote repository's URL. 
So, to push your changes to the remote repository, 
you could've used either the command: git push git@github.com:git/git.git yourbranchname or git push origin yourbranchname


## Pull Request
A pull request (or PR) is a way to alert a repo's owners that you want to make some changes to their code. 
It allows them to review the code and make sure it looks good before putting your changes on the master branch.


## Get Changes on Github
`git pull origin master`

check all new commits: `git log`


## View Differences
1) run: `git diff`


## Remove a Branch
### Locally
`git branch -d <branch_name>`

### Remote
`git push <remote_name> --delete <branch_name>`

## Remove tracked file/directory
### File
`git rm --cached <file>`

### Directory
`git rm --cahced -r dir/`


---

# Vim

## .vimrc
```{}
syntax on

" Color scheme


" New lines inherit indentation of previous line
set autoindent
" Convert tabs to spaces
set expandtab

" Search highlighting
set hlsearch
" Ignore case when searching
set ignorecase

" Highlight line currently under cursor
set cursorline
" Line numbers on sidebar
set number

" Spellchecking
set spell
```

---

# psql
[Cheat sheet](https://quickref.me/postgres)
```{r psql, echo=F}
t4 <- read_excel(f, sheet="psql-commands")
# Show all rows in table (no pagination)
datatable(t4, options=list(paging=FALSE), rownames=F)
```


## Export table to CSV
- `\copy table TO '<path>' CSV`
- `\copy table(col1,col1) TO '<path>' CSV`
- `\copy (SELECT...) TO '<path>' CSV`


## Backup
Use pg_dumpall to backup all databases

`$ pg_dumpall -U postgres > all.sql`

Use pg_dump to backup a database

`$ pg_dump -d mydb -f mydb_backup.sql`

- &nbsp; `-a` &nbsp; Dump only the data, not the schema
- &nbsp; `-s` &nbsp; Dump only the schema, no data
- &nbsp; `-c` &nbsp; Drop database before recreating
- &nbsp; `-C` &nbsp; Create database before restoring
- &nbsp; `-t` &nbsp; Dump the named table(s) only
- &nbsp; `-F` &nbsp; Format (`c`: custom, `d`: directory, `t`: tar)

Use `pg_dump -?` to get the full list of options


## Restore

### psql

`$ psql -U user mydb < mydb_backup.sql`

### pg_restore

`$ pg_restore -d mydb mydb_backup.sql -c`

- `-U   Specify a database user`
- `-c   Drop database before recreating`
- `-C   Create database before restoring`
- `-e   Exit if an error has encountered`
- `-F   Format (c: custom, d: directory, t: tar, p: plain text sql(default))`

Use pg_restore -? to get the full list of options
