---
title: "Health"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, error = T)

# https://readxl.tidyverse.org/
library(tidyverse)
library(readxl)
library(DT)
```

---

# Workout Routines

The following regimen outlines a typical week of lifts. Non-lift days should be supplemented with active recovery workouts.An example week could look something like this:

1. **Monday**: Back, Biceps, Abs
1. **Tuesday**: Chest & Triceps
1. **Wednesday**: Run
1. **Thursday**: Legs
1. **Friday**: Shoulders & Traps
1. **Saturday**: Long Walk
1. **Sunday**: Run

```{r, include=F}
workouts <- "./static/workouts.xlsx"
w1 <- read_excel(workouts, sheet=1)
w2 <- read_excel(workouts, sheet=2)
w3 <- read_excel(workouts, sheet=3)
w4 <- read_excel(workouts, sheet=4)
```

## Back, Biceps, Abs

```{r, eval=T}
datatable(w1, options=list(scrollX=T, dom='t'))
```


## Chest & Triceps

```{r, eval=T}
datatable(w2, options=list(scrollX=T, dom='t'))
```

## Legs

```{r, eval=T}
datatable(w3, options=list(scrollX=T, dom='t'))
```

## Shoulders & Traps

```{r, eval=T}
datatable(w4, options=list(scrollX=T, dom='t'))
```

